---
title: "Buoy"
author: "Yibing Wang"
date: "2024-09-26"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(lubridate)
library(dplyr)
library(ggplot2)
library(corrplot)
```

```{r}
file_root <- "https://www.ndbc.noaa.gov/view_text_file.php?filename=44013h"
tail <- ".txt.gz&dir=data/historical/stdmet/"

years <- 1985:2023

all_data <- list()

# Loop Through Each Year
for (year in years) {
  path <- paste0(file_root, year, tail)
  # Skip line
  temp_lines <- readLines(path, n = 2)
  skip_lines <- if (grepl("YYYY", temp_lines[2])) 1 else 2
  
  # Read Header
  header <- tryCatch({
    scan(path, what = 'character', nlines = 1, quiet = TRUE)
  }, error = function(e) NULL)
  
  # if the Header is Successfully Read
  if (!is.null(header)) {
    buoy_data <- fread(path, header = FALSE, skip = skip_lines, fill = Inf)
    # Adjust Column Names Based on the Header
    if (ncol(buoy_data) > length(header)) {
      setnames(buoy_data, 1:length(header), header)
      additional_cols <- ncol(buoy_data) - length(header)
      if (additional_cols > 0) {
        # Name Extra Columns as V1, V2, V3.
        setnames(buoy_data, (length(header) + 1):ncol(buoy_data), paste0("V", 1:additional_cols))
      }
    } else {
      setnames(buoy_data, header)
    }
    
    # date to integers
    date_cols <- intersect(c("YY", "YYYY", "#YY", "MM", "DD", "hh"), names(buoy_data))
    buoy_data[, (date_cols) := lapply(.SD, as.integer), .SDcols = date_cols]
    
    if (all(c("YY", "MM", "DD", "hh") %in% colnames(buoy_data))) {
      # Adjust to YYYY
      buoy_data[, date := make_datetime(1900 + YY, MM, DD, hh)]
    } else if (all(c("YYYY", "MM", "DD", "hh") %in% colnames(buoy_data))) {
      buoy_data[, date := make_datetime(YYYY, MM, DD, hh)]
    }
    
    #"TIDE" 
    buoy_data[, TIDE := NA_real_]
    
    # Store the Data Table in the List
    all_data[[as.character(year)]] <- buoy_data
  }
}

# Combine All Data
full_dataset <- rbindlist(all_data, fill = TRUE)

# to YYYY
if ("YY" %in% names(full_dataset)) {
  full_dataset[, YY := as.integer(YY)]
  full_dataset[, YY := ifelse(!is.na(YY), 1900 + YY, YY)]
}

# Merge into 'year' Column
if ("#YY" %in% names(full_dataset)) {
  full_dataset[, `#YY` := as.numeric(sub("^#", "", `#YY`))]
}

full_dataset[, year := fcoalesce(as.numeric(YYYY),
                                 as.numeric(YY),
                                 as.numeric(`#YY`))]

# Remove YYYY, YY, #YY
full_dataset[, c("YYYY", "YY", "#YY") := NULL]

# Reorder 'year' at the Beginning
setcolorder(full_dataset, c("year", setdiff(names(full_dataset), "year")))

# 'WDIR'
full_dataset[, WDIR := fcoalesce(as.numeric(WDIR), as.numeric(WD))]
full_dataset[, WD := NULL]
full_dataset[, PRES := fcoalesce(as.numeric(PRES), as.numeric(BAR))]
full_dataset[, BAR := NULL]

print(head(full_dataset))
```
# Part B

```{r}
# List of variables to check for placeholder values
variables_to_check <- c("WDIR", "WSPD", "GST", "WVHT", "DPD", "APD", "MWD", "PRES", "ATMP", "WTMP", "DEWP", "VIS")

# replace 999 with NA
replace_placeholders <- function(x) {
  x_numeric <- as.numeric(x)
  x_numeric[x_numeric %in% c(999, 9999, 999.0, 9999.0)] <- NA
  return(x_numeric)
}

for (var in variables_to_check) {
  if (var %in% names(full_dataset)) {
    full_dataset[, (var) := replace_placeholders(get(var))]
  }
}

print(head(full_dataset))
```

```{r}
# Add 'date' column; 'date' format; number of NAs per variable
if (!"date" %in% names(full_dataset)) {
  full_dataset[, date := make_datetime(year, MM, DD, hh)]
}
full_dataset[, date := as.POSIXct(date)]

na_counts <- sapply(full_dataset[, variables_to_check, with = FALSE], function(x) sum(is.na(x)))
print(na_counts)
```

```{r}
na_by_year <- full_dataset[, lapply(.SD, function(x) sum(is.na(x))), by = year, .SDcols = variables_to_check]

na_by_year_melted <- melt(na_by_year, id.vars = "year", variable.name = "Variable", value.name = "NA_Count")
ggplot(na_by_year_melted, aes(x = factor(year), y = NA_Count, fill = Variable)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Variable, scales = "free_y") +
  labs(title = "Number of Missing Values per Variable by Year",
       x = "Year",
       y = "Count of Missing Values") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
ggplot(na_by_year_melted, aes(x = factor(year), y = NA_Count, fill = Variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Number of Missing Values per Variable by Year",
       x = "Year",
       y = "Count of Missing Values",
       fill = "Variable") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "bottom")
```
I think it is important to convert missing data to NA, but it is not always appropriate. 
converting to NA can help us using r function better since it is the standard representation for missing value in r. 
However, in some cases, if we convert the data to NA without fully understand the meaning, it might lead to misinterpretation of the data. 

# C

```{r}
temperature_vars <- c("ATMP", "WTMP")
for (var in temperature_vars) {
  if (var %in% names(full_dataset)) {
    full_dataset[, (var) := as.numeric(get(var))]
  }
}
atmp_data <- full_dataset[!is.na(ATMP) & year <= 2006]
wtmp_data <- full_dataset[!is.na(WTMP)]
```

```{r}
# annual mean ATMP
annual_atmp <- atmp_data[, .(
  Mean_ATMP = mean(ATMP, na.rm = TRUE)
), by = year]
```

```{r}
# annual mean WTMP
annual_wtmp <- wtmp_data[, .(
  Mean_WTMP = mean(WTMP, na.rm = TRUE)
), by = year]
```

```{r}
ggplot(annual_atmp, aes(x = year, y = Mean_ATMP)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(title = "Annual Average Air Temperature (ATMP)",
       x = "Year",
       y = "Mean Air Temperature (°C)") +
  theme_minimal()
```

```{r}
ggplot(annual_wtmp, aes(x = year, y = Mean_WTMP)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  labs(title = "Annual Average Water Temperature (WTMP)",
       x = "Year",
       y = "Mean Water Temperature (°C)") +
  theme_minimal()
```

```{r}
#up to 2006
if (nrow(annual_atmp) >= 3) {  # Ensure enough data points
  model_ATMP <- lm(Mean_ATMP ~ year, data = annual_atmp)
  summary_ATMP <- summary(model_ATMP)
  print(summary_ATMP)
  
  rate_ATMP_per_decade <- coef(model_ATMP)["year"] * 10
  cat("Air Temperature is changing at a rate of ", round(rate_ATMP_per_decade, 4), "°C per decade (up to 2006).\n")
  plot(model_ATMP)
} else {
  cat("Not enough data points for ATMP to perform regression analysis.\n")
}
```

```{r}
if (nrow(annual_wtmp) >= 3) {
  model_WTMP <- lm(Mean_WTMP ~ year, data = annual_wtmp)
  summary_WTMP <- summary(model_WTMP)
  print(summary_WTMP)
  
  rate_WTMP_per_decade <- coef(model_WTMP)["year"] * 10
  cat("Water Temperature is changing at a rate of ", round(rate_WTMP_per_decade, 4), "°C per decade.\n")
  plot(model_WTMP)
} else {
  cat("Not enough data points for WTMP to perform regression analysis.\n")
}
```

# D

```{r}
rainfall_data <- read.csv("Rainfall.csv")
summary(rainfall_data)
```

```{r}
rainfall_data$DATE <- as.POSIXct(rainfall_data$DATE, format="%Y%m%d %H:%M")
rainfall_data <- rainfall_data %>%
  mutate(
    year = format(DATE, "%Y"),
    MM = format(DATE, "%m"),
    DD = format(DATE, "%d"),
    HH = format(DATE, "%H")
  )
```

```{r}
# rainfall(HPCP)
rainfall_summary <- rainfall_data %>%
  summarize(
    count = n(),
    mean = mean(HPCP, na.rm = TRUE),
    std = sd(HPCP, na.rm = TRUE),
    min = min(HPCP, na.rm = TRUE),
    Q1 = quantile(HPCP, 0.25, na.rm = TRUE),
    median = median(HPCP, na.rm = TRUE),
    Q3 = quantile(HPCP, 0.75, na.rm = TRUE),
    max = max(HPCP, na.rm = TRUE)
  )

print(rainfall_summary)
```

```{r}
ggplot(rainfall_data, aes(x = HPCP)) +
  geom_histogram(binwidth = 0.01, fill = "blue", color = "black") +
  labs(title = "Distribution of Rainfall", x = "Rainfall", y = "Frequency")
```


```{r}
total_precipitation <- rainfall_data %>%
  group_by(year) %>%
  summarize(total_precipitation = sum(HPCP, na.rm = TRUE))

ggplot(total_precipitation, aes(x = as.numeric(year), y = total_precipitation)) +
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(title = "Total Annual Precipitation", x = "Year", y = "Total Precipitation (inches)") +
  theme_minimal()
```

```{r}
average_precipitation <- rainfall_data %>%
  group_by(year) %>%
  summarize(average_precipitation = mean(HPCP, na.rm = TRUE))

ggplot(average_precipitation, aes(x = as.numeric(year), y = average_precipitation)) +
  geom_line(color = "green") +
  geom_point(color = "orange") +
  labs(title = "Average Annual Precipitation", x = "Year", y = "Average Precipitation") +
  theme_minimal()
```

```{r}
# total precipitation for each month
monthly_precipitation <- rainfall_data %>%
  group_by(year, MM) %>%
  summarize(total_monthly_precipitation = sum(HPCP, na.rm = TRUE))
```

```{r}
ggplot(monthly_precipitation, aes(x = MM, y = total_monthly_precipitation)) +
  geom_boxplot(aes(group = MM), fill = "lightblue") +
  facet_wrap(~ year) +
  labs(title = "Distribution of Monthly Precipitation for Each Year",
       x = "Month",
       y = "Total Precipitation (inches)") +
  theme_minimal()
```

```{r}
#buoy
atmp_data <- full_dataset[!is.na(ATMP) & year <= 2006]

monthly_atmp <- atmp_data[, .(
  Mean_ATMP = mean(ATMP, na.rm = TRUE)
), by = .(year, MM)]

monthly_atmp[, date := as.Date(paste(year, MM, "01", sep = "-"))]

wtmp_data <- full_dataset[!is.na(WTMP) & year <= 2006]

monthly_wtmp <- wtmp_data[, .(
  Mean_WTMP = mean(WTMP, na.rm = TRUE)
), by = .(year, MM)]

monthly_wtmp[, date := as.Date(paste(year, MM, "01", sep = "-"))]
```

```{r}
ggplot(monthly_atmp, aes(x = date, y = Mean_ATMP)) +
  geom_line(color = "blue") +
  labs(title = "Monthly Average Air Temperature (ATMP) up to 2006",
       x = "Date",
       y = "Mean Air Temperature (°C)") +
  theme_minimal()
```

```{r}
ggplot(monthly_wtmp, aes(x = date, y = Mean_WTMP)) +
  geom_line(color = "red") +
  labs(title = "Monthly Average Water Temperature (WTMP)",
       x = "Date",
       y = "Mean Water Temperature (°C)") +
  theme_minimal()
```

```{r}
#ATMP
ggplot(atmp_data, aes(x = factor(MM), y = ATMP)) +
  geom_boxplot(aes(group = MM), fill = "lightblue") +
  facet_wrap(~ year, ncol = 4) +
  labs(title = "Distribution of Monthly Air Temperature (ATMP) for Each Year up to 2006",
       x = "Month",
       y = "Air Temperature (°C)") +
  theme_minimal()
```

```{r}
#WTMP
ggplot(wtmp_data, aes(x = factor(MM), y = WTMP)) +
  geom_boxplot(aes(group = MM), fill = "lightblue") +
  facet_wrap(~ year, ncol = 4) +
  labs(title = "Distribution of Monthly Water Temperature (WTMP) for Each Year",
       x = "Month",
       y = "Water Temperature (°C)") +
  theme_minimal()
```

```{r}
monthly_precipitation$year <- as.numeric(monthly_precipitation$year)
monthly_precipitation$MM <- as.numeric(monthly_precipitation$MM)

merged_data <- monthly_atmp %>%
  inner_join(monthly_wtmp, by = c("year", "MM")) %>%
  inner_join(monthly_precipitation, by = c("year", "MM"))
```

```{r}
#correlation
selected_data <- merged_data %>%
  select(Mean_ATMP, Mean_WTMP, total_monthly_precipitation)

#correlation matrix
correlation_matrix <- cor(selected_data, use = "complete.obs")

corrplot::corrplot(correlation_matrix, method = "circle", type = "upper", 
                   tl.col = "black", title = "Correlation Matrix", mar = c(0, 0, 2, 0))
```
Strong correlation between air and water temperatures, but weak negative correlation with precipitation


```{r}
model <- lm(total_monthly_precipitation ~ Mean_ATMP + Mean_WTMP, data = merged_data)

summary(model)
plot(model)
```
Just like what the correlation matrix tells me, a linear model won't be a good idea to use for finding the relationship between Buoy data and precipitation.


```

